Robust Winner Determination and Simultaneous Elicitation of Scoring Rules and Preferences

### 2019-04-09 draft.tex
beta construction

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ about the preference of the agent $j$. 
The adversary's goal is to make the score of $y$ as high as possible and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \pprefeq_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \pprefeq_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \pprefeq_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \pprefeq_j y) ∨ (x \ppref_j a)).
\end{split}
\end{align} 
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
\end{claim}
It follows that $\hat{v}_{j}(x)$, the rank of $x$ in the PMR-maximizing linear order of agent $j$, 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\pprefeq_j}(x)}$, where ${\pprefeq_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\pprefeq_j$ (thus including $x$ itself), and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\beta$, where $\beta = \card{{\pinc_j}(x)}$ if $(x \pprefeq_j y)$ and $\beta = 0$ otherwise and with ${\pinc_j}(x)$ designating the alternatives incomparable with $x$ in $\pprefeq_j$ \commentBN{Still not good. In the set of alternatives incomparable with $x$ there could be alternatives strictly better than $y$. It is exactly the problem we had before. Instead, $\beta$ represents the number of alternatives incomparable with $x$ that are no better than $y$: $\beta = \card{A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))}$ }.

>>>>

We are interested in computing $\PMR(x, y; \pprofile, W)$. This represents the “worst” difference of score, thus the difference of score between $y$ and $x$ under the worst case preferences compatible with $\pprofile$ and $W$, where the worst case is the one that maximizes this difference of score.
Consider our knowledge $\ppref_j$ about the preference of the agent $j$. 
The adversary's goal is to make the score of $y$ as high as possible and the score of $x$ as low as possible. 
To do this, he should complete $\ppref_j$ to $\pref_j$ by putting above $x$ as many alternatives as he can, that is, all the alternatives except those that are known to be worse than $x$ (those $a$ such that $x \pprefeq_j a$); and similarly, put below $y$ all the alternatives he can, thus, put above $y$ only those $a$ such that $a \ppref_j y$. 
Sometimes these two objectives conflict: when an alternative should go above $x$ according to the first objective and below $y$ according to the second one (because $¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y)$), and $x$ is known to be better than $y$ (thus $x \pprefeq_j y$). 
In this case, the first objective should take priority, and $a$ should be put above $x$, which will move both $x$ and $y$ one rank lower than if $a$ had been put below $y$. 
This maximizes the adversary’s interests: because the weight vector is convex, the difference of scores will be lower when both alternatives are ranked lower (Equation \ref{eq:convexity}), and that difference of scores is in favor of $x$ when $x \ppref_j y$, thus to be minimized from the point of view of the adversary.
To summarize:
\begin{align} 
\label{eq:compl}
\begin{split}
a \pref_j x &⇔ ¬(x \pprefeq_j a)\\
y \pref_j a &⇔ ¬(a \ppref_j y) ∧ (¬(x \pprefeq_j y) ∨ (x \pprefeq_j a)).
\end{split}
\end{align} 
\commentBN{I think it would be more intuitive writing: \\ $ y \pref_j a ⇔ ¬(a \ppref_j y) ∧ ¬((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))$}
We just sketched a proof of the following claim, and leave the rest of the proof to the reader.
\begin{claim}
There exists a completion $\hat{\profile} \in C(\pprofile)$ such that $\PMR(x,y; \pprofile, W) = \max_{\w \in W} [ s(y; \hat{\profile}, \w) - s(x; \hat{\profile}, \w) ]$ and such that the linear order $\hat{v}_{j}$ of each agent $j$ satisfies (\ref{eq:compl}).
\end{claim}
It follows that $\hat{v}_{j}(x)$, the rank of $x$ in the PMR-maximizing linear order of agent $j$, 
is $\hat{v}_{j}(x) = 1+\card{A}-\card{{\pprefeq_j}(x)}$, where ${\pprefeq_j}(x)$ designates the set of alternatives that are less good than $x$ according to $\pprefeq_j$ (thus including $x$ itself), and the rank of $y$ is $\hat{v}_{j}(y)=1+\card{{\pprefinv_j}(y)}+\card{\beta}$, where $\card{\beta} = \card{A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))}$ if $(x \pprefeq_j y)$ and $\card{\beta} = 0$ otherwise.
In the case $(x \pprefeq_j y)$, intuitively $\beta$ represents the number of alternatives incomparable with $x$ that are no better than $y$. Let's reason about what alternatives must be placed above $y$ by negating the second equation of \ref{eq:compl}:
\begin{align}
\label{eq:betaproof} 
\begin{split}
a \prefeq_j y &⇔ (a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))\\
a \pref_j y &⇔ (a \neq_j y) ∧ [(a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a))]
\end{split}
\end{align} 
But the term in the square brackets implies $(a \neq_j y)$, indeed:
\[(a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a)) \Rightarrow (a \neq_j y)\]
$(a \ppref_j y) \Rightarrow (a \neq_j y)$ is always true, and we can rewrite $¬(x \pprefeq_j a)$ as $(a \ppref_j x) ∨ (a \pinc_j x)$, where $\pinc_j$ designates incomparability. Thus:
\begin{gather*} 
(x \pprefeq_j y) ∧ ¬(x \pprefeq_j a) \Rightarrow (a \neq_j y) \\
(x \pprefeq_j y) ∧ ((a \ppref_j x) ∨ (a \pinc_j x)) \Rightarrow (a \neq_j y) \\
((x \pprefeq_j y) ∧ (a \ppref_j x)) ∨ ((x \pprefeq_j y) ∧ (a \pinc_j x)) \Rightarrow (a \neq_j y).
\end{gather*}
So we can go back to \ref{eq:betaproof} dropping the $(a \neq_j y)$ term:
\[a \pref_j y ⇔ (a \ppref_j y) ∨ ((x \pprefeq_j y) ∧ ¬(x \pprefeq_j a) ∧ ¬(a \ppref_j y))\]
In case $(x \pprefeq_j y)$, we say an alternative $a \in \beta$ iff $¬((x \pprefeq_j a) ∨ (a \ppref_j y))$. Thus defining $\beta=A \setminus ({\pprefeq_j}(x) \cup {\pprefinv_j}(y))$.
